Proof nets for first-order additive linear logic

### 2018-03-23 Archive/ALL1-backup-8.tex
some cleanup

\newcommand\subs[1]{\textsc{sub}(#1)}
\newcommand\dom[1]{\textsc{dom}(#1)}

@@

First-order terms and the formulas of first-order $\all$ are generated by the following grammars.
%
\setMidspace{5pt}
\[
\begin{array}{@{}l@{}l}
	t &\Coloneqq a \Mid x \Mid f(t_1,\dots,t_n)
\\[10pt]
	P &\Coloneqq p(t_1,\dots,t_n) \Mid \dual p(t_1,\dots,t_n)
\\[10pt]
	A &\Coloneqq P \Mid A\+A \Mid A\*A \Mid \exists x.A \Mid \forall a.A
\end{array}
\]
%
Negation $(\dual{\,\cdot\,})$ is applied to predicate symbols, $\dual p$, as a matter of convenience. The \defn{dual} $\dual A$ of an arbitrary formula $A$ is given by DeMorgan. We use the following notational conventions.
%
\[
\begin{tabular}{@{}lll@{}}
	$a,b,c$ & $\in \varA$ 		& universally quantified variables (and free variables)\\
	$x,y,z$ & $\in \varE$		& existentially quantified variables \\
	$f,g,h$ & $\in \Sigma_f$	& $n$-ary $(n\geq 0)$ function symbols from a fixed alphabet $\Sigma_f$\\
	$p,q,r$ & $\in \Sigma_p$	& $n$-ary $(n\geq 0)$ predicate symbols from a fixed alphabet $\Sigma_p$ \\
	$s,t,u$ & $\in \terms$ 		& first-order terms over $\varA$, $\varE$, $\Sigma_f$, and $\Sigma_p$ \\
	$P,Q,R$ & $\in \atom$		& atomic propositions \\
	$A,B,C$ & $\in \form$		& \all1 formulas \\
\end{tabular}
\]
%
A \defn{sequent} $\seq AB$ is a pair of formulas $A$ and $B$. A sequent calculus for \all1 is given in Figure~\ref{fig:sequent calculus}, where each rule has a symmetric counterpart that applies to the first formula in the sequent. We write $\prf\pi AB$ for a proof $\pi$ with conclusion sequent $\seq AB$. Anticipating the requirements of proof nets, we will assume the following convention for \all1 proofs.

@@

A \defn{witness map} $\sigma\colon \varE \rightharpoonup \terms$ is a substitution map which assigns terms to existential variables, given as a (finite) partial function. The \defn{empty} witness map $\varnothing$ is the everywhere undefined partial function, and the map $\sigma[t/x]$ substitutes $t$ for $x$ and is as $\sigma$ everywhere else. We express by $\sigma(x)=\bot$ that $\sigma$ is undefined for $x$. The map $\sigma\minus x$ is undefined on $x$ and otherwise as $\sigma$.

We will write $A\sigma$ for the application of the witness map $\sigma\colon\varE\rightharpoonup\terms$ to the formula $A$ as a substitution map. It is applied to a sequent proof, $\pi\sigma$, by applying it to each formula in the proof $\pi$. The \defn{composition} of two maps is written $\sigma\tau$, where $A(\sigma\tau)=(A\sigma)\tau$. We define the following further operations on witness maps.

\[
\begin{tabular}{@{}lp{.75\textwidth}@{}}
	$\sigma\gen\tau$
&
	A witness map is \defn{more general} than another, 
	$\sigma\gen\tau$, if there is a map $\rho$ such that 
	$\sigma\rho=\tau$.
\\[5pt]	
	$\sigma\coh\tau$
&
	Two witness maps are \defn{coherent}, $\sigma\coh\tau$,
	if there is a map $\rho$ such that $\sigma\rho=\tau\rho$.
\\[5pt]
	$\sigma\join\tau$
&
	The \defn{join} of two witness maps $\sigma\join\tau$ is defined
	if they are coherent, and is the least map $\rho$ such that
	$\sigma\gen\rho$ and $\tau\gen\rho$.
\end{tabular}%
\]

A \defn{witness linking} $\lambda_\Sigma$ is a linking $\lambda$ with a \defn{witness labelling}  $\Sigma\colon\lambda\to \varE\rightharpoonup\terms$ that assigns each link $(C,D)$ a witness map. 
We may use and define $\lambda_\Sigma$ as a set of \defn{witness links} $\link CD$, which indicates the link $(C,D)$ in $\lambda$ where $\Sigma(C,D)=\sigma$.

\begin{definition}
A \defn{witness pre-net} $\net{\lambda_\Sigma}AB$ is a pre-net with a witness linking $\lambda_\Sigma$.
\end{definition}

We assume the existential variables occurring in the range of $\Sigma$ to be fresh, i.e.\ distinct from any existentially quantified variable in $\seq AB$. We abbreviate by $a\in \sigma$ the statement that a (universal) variable $a$ occurs free in the range of $\sigma$, i.e.\ that $a\in\fv(\sigma(x))$ for some $x$.

@@

To an axiom link $(P,Q)$ we assign an \defn{initial witness map} $\sigma=\init PQ$ as the least witness map in $(\gen)$ such that $P\sigma=\dual Q\sigma$ --- in other words, $\init PQ$ is the most general unifier ($\textsc{mgu}$) of $P$ and $\dual Q$. An axiom linking $\lambda$ with the initial witness map assigned to each link will be written $\lambda_\star$:
\[
	\lambda_\star = \{~\link PQ~\mid~(P,Q)\in\lambda~,~\sigma=\init PQ~\}~.
\]

For two witness labellings $\Sigma$ and $\Xi$ on the same linking $\lambda$, one is \defn{more general} than the other, $\Sigma\gen\Xi$, if $\Sigma(C,D)\gen\Xi(C,D)$ for every link $(C,D)\in\lambda$.  $\lambda_\Sigma\gen\lambda_\Xi$,

Observe that for an axiom linking $\lambda$, the initial linking $\lambda_\star$ is the most general among the witness linkings $\lambda_\Sigma$ for any $\Sigma$.

@@

\noindent
A pre-net $\net{\lambda_\Sigma}AB$ \defn{strict-coalesces} if it reduces in $(\scoal)$ to a single link $\net{\{\link[\varnothing]AB\}}AB$.
\end{definition}

To be explicit, to generate coalescence as a rewrite relation means the following: a pre-net coalesces to another, $\net{\lambda_\Sigma} AB\scoal \net{{\lambda'_{\Sigma'}}}AB$ if and only if % there is a rewrite rule $\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD$ such that $C\leq A$, $D\leq B$, ${\kappa_\Xi}\subseteq{\lambda_\Sigma}$, $\kappa'_{\Xi'}\subseteq{\lambda'_{\Sigma'}}$, and ${\lambda_\Sigma}-{\kappa_\Xi}={\lambda'_{\Sigma'}}-\kappa'_{\Xi'}$.
there is a rewrite rule
\[
	\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD
\] 
such that 
\[
	C\leq A \qquad
	D\leq B \qquad
	\kappa_\Xi\subseteq\lambda_\Sigma \qquad
	\kappa'_{\Xi'}\subseteq\lambda'_{\Sigma'} \quad
	\text{and} \quad
	\lambda_\Sigma-\kappa_\Xi=\lambda'_{\Sigma'}-\kappa'_{\Xi'}~.
\]

@@

%\end{document}

A link is \defn{strict} if its domain consists of exactly the existentially quantified variables that its formulas are in scope of. That is, a link $\link CD$ in a pre-net $\net\lambda AB$ is strict if
\[
	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
\]
A linking or pre-net is \defn{strict} if all its links are.

For a labelled linking $\lambda$ and an existential variable $x$, write $\lambda(x)\subseteq\terms$ for the \defn{witness set} of $x$, which collects the terms assigned to $x$ by the links in $\lambda$:
\[
	\lambda(x) = \{~t~\mid~\sigma(x) = t~,~\link AB\in\lambda~\}~.
\]
%
Given a labelled linking $\lambda$, a \defn{slice} $S$ of a formula $A$ is a set of formulas given by: $S=\{A\}\cup S'$ where $S'$ is as follows; if $A$ is of the form
%
\begin{itemize}
	\item
\makebox[40pt][l]{$P$:}				$S'=\varnothing$.
	\item
\makebox[40pt][l]{$B\+C$:} 			$S'=S_B\cup S_C$ where $S_B$ is a slice of $B$ and $S_C$ a slice of $C$.
	\item
\makebox[40pt][l]{$B\*C$:}			$S'$ is a slice of $B$ or a slice of $C$.
	\item
\makebox[40pt][l]{$\exists x.B$:}	$S'$ is a slice of $B$ if the quantifier is vacuous, $x\notin\fv(B)$, and
	\\[\itemsep]
	\hspace*{40pt}					$S'=\uplus_{t\,\in\,\lambda(x)}S_t$ where each $S_t$ is a slice of $B[t/x]$ otherwise.
	\item
\makebox[40pt][l]{$\forall a.B$:}	$S'$ is a slice of $B$.
\end{itemize}
%
A \defn{slice} of a labelled pre-net $\net\lambda AB$ is a set of links
\[
	\{~\link CD\in \lambda~\mid~(C\sigma,D\sigma)\in S_A\times S_B~\}
\]
where $S_A$ is a slice of $A$ and $S_B$ a slice of $B$. A labelled pre-net is \defn{correct} if every slice is a singleton. A coherent pre-net $\net\lambda AB$ is \defn{correct} if its initial labelling $\net{\lambda_\star}AB$ is. Observe that for a \emph{labelled} pre-net correctness and compatibility are independent, while a non-labelled pre-net must be coherent to be correct, since $\lambda_\star$ is otherwise not defined.

\begin{definition}
An \all1 \defn{proof net} is a correct pre-net. An \all1 \defn{labelled proof net} is a correct and coherent labelled pre-net.
\end{definition}

\begin{lemma}
If a labelled pre-net coalesces, it is correct. 
\end{lemma}

\begin{proof}
By induction on the coalescence path we will show that if $\net\lambda AB$ coalesces to $\net{\{\link AB\}}AB$ then $\net{\{\link[\varnothing]{A\sigma}{B\sigma}\}}{A\sigma}{B\sigma}$ is correct. The statement is then the special case $\sigma=\varnothing$.

The result $\net{\{\link[\varnothing]{A\sigma}{B\sigma}\}}{A\sigma}{B\sigma}$ is correct: its only slice is $\{\link[\varnothing]{A\sigma}{B\sigma}\}$, since $A\sigma$ is in any slice $S_{A\sigma}$ of $A\sigma$, and $B\sigma$ is in any slice $S_{B\sigma}$ of $B\sigma$.

\renewcommand{\+}{{+}}
\renewcommand{\*}{{\times}}
\begin{align}
	\net{\{\link AB,\link[\tau]AC\}}A{B\*C}
& ~\scoal~ 
	\net{\{\link[\sigma\join\tau]A{B\*C}\}}A{B\*C}\quad\text{if }\sigma\coh\tau
\tag{$\*$}
\end{align}

Let $\rho=\sigma\join\tau$, and assume that $\net{\{\link[\varnothing]{A\rho}{B\rho\*C\rho}\}}{A\rho}{B\rho\*C\rho}$ is correct.

@@

\\ \\
    \vc{\infer{\crule[\sigma\minus x]A{\exists x.B}}{\crule AB}}
&
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\exists x.B}
    	\Vlink[red,label={~$\scriptstyle\sigma$}]{1,4}
    \end{tikzpicture}}
&\rightsquigarrow&
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\exists x.B}
    	\Vlink[red,label={~$\scriptstyle{\sigma\!\minus\!x}$}]{1,1}
    \end{tikzpicture}}
\\ \\
    \vc{\infer[\!\!\scriptstyle{a\,\notin\,\fv(A\sigma)}]{\crule A{\forall a.B}}{\crule AB}}
&    
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\forall a.B}
    	\Vlink[red,label={~$\scriptstyle\sigma$}]{1,4}
    \end{tikzpicture}}
&\stackrel{a\,\notin\,\fv(A\sigma)}\rightsquigarrow&
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\forall a.B}
    	\Vlink[red,label={~$\scriptstyle\sigma$}]{1,1}
    \end{tikzpicture}}
\end{array}
\]
\par\bigskip\hrule
\caption{Coalescence derivation rules}
\label{fig:coalescence}
\end{figure}

>>>>

\newcommand\subs[1]{\textsc{sub}(#1)}
\newcommand\poss[1]{\textsc{pos}(#1)}
\newcommand\dom[1]{\textsc{dom}(#1)}

@@

First-order terms and the formulas of first-order $\all$ are generated by the following grammars.
%
\setMidspace{5pt}
\[
\begin{array}{@{}l@{}l}
	t &\Coloneqq a \Mid x \Mid f(t_1,\dots,t_n)
\\[10pt]
	P &\Coloneqq p(t_1,\dots,t_n) \Mid \dual p(t_1,\dots,t_n)
\\[10pt]
	A &\Coloneqq P \Mid A\+A \Mid A\*A \Mid \exists x.A \Mid \forall a.A
\end{array}
\]
%
Negation $(\dual{\,\cdot\,})$ is applied to predicate symbols, $\dual p$, as a matter of convenience. The \defn{dual} $\dual A$ of an arbitrary formula $A$ is given by DeMorgan. We use the following notational conventions.
%
\[
\begin{tabular}{@{}lll@{}}
	$a,b,c$ & $\in \varA$ 		& universally quantified variables (and free variables)\\
	$x,y,z$ & $\in \varE$		& existentially quantified variables (and term-language variables)\\
	$f,g,h$ & $\in \Sigma_f$	& $n$-ary $(n\geq 0)$ function symbols from a fixed alphabet $\Sigma_f$\\
	$p,q,r$ & $\in \Sigma_p$	& $n$-ary $(n\geq 0)$ predicate symbols from a fixed alphabet $\Sigma_p$ \\
	$s,t,u$ & $\in \terms$ 		& first-order terms over $\varA$, $\varE$, $\Sigma_f$, and $\Sigma_p$ \\
	$P,Q,R$ & $\in \atom$		& atomic propositions \\
	$A,B,C$ & $\in \form$		& \all1 formulas \\
\end{tabular}
\]
%
A \defn{sequent} $\seq AB$ is a pair of formulas $A$ and $B$. A sequent calculus for \all1 is given in Figure~\ref{fig:sequent calculus}, where each rule has a symmetric counterpart that applies to the first formula in the sequent. We write $\prf\pi AB$ for a proof $\pi$ with conclusion sequent $\seq AB$. Anticipating the requirements of proof nets, we will assume the following convention for \all1 proofs.

@@

A \defn{witness map} $\sigma\colon \varE \rightharpoonup \terms$ is a substitution map which assigns terms to existential variables, given as a (finite) partial function. Its \defn{domain} $\dom(\sigma)$ are those variables for which it is defined. The \defn{empty} witness map $\varnothing$ is the everywhere undefined partial function, and the map $\sigma[t/x]$ substitutes $t$ for $x$ and is as $\sigma$ everywhere else. We express by $\sigma(x)=\bot$ that $\sigma$ is undefined for $x$. The map $\sigma\minus x$ is undefined on $x$ and otherwise as $\sigma$.

We will write $A\sigma$ for the application of the witness map $\sigma\colon\varE\rightharpoonup\terms$ to the formula $A$ as a substitution map. It is applied to a sequent proof, $\pi\sigma$, by applying it to each formula in the proof $\pi$. The \defn{composition} of two maps is written $\sigma\tau$, where $A(\sigma\tau)=(A\sigma)\tau$.

A \defn{witness linking} $\lambda_\Sigma$ is a linking $\lambda$ with a \defn{witness labelling} $\Sigma\colon\lambda\to \varE\rightharpoonup\terms$ that assigns each link $(C,D)$ a witness map. We may use and define $\lambda_\Sigma$ as a set of \defn{witness links} $\link CD$, where $(C,D)$ in $\lambda$ and $\Sigma(C,D)=\sigma$.

\begin{definition}
A \defn{witness pre-net} $\net{\lambda_\Sigma}AB$ is a sequent $\seq AB$ with a witness linking $\lambda_\Sigma$.
%
%\begin{description}
%	\item[Exact coverage]
%The domain of $\sigma$ is exactly the variables existentially quantified in $\seq AB$ that will be bound in $C$ and $D$, if they occur:
%\[
%	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
%			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
%\]
%
%	\item[Eigenvariables not free]
%A universal variable $a\in\varA$ quantified in $C$ or $D$ may not occur free in the range of $\sigma$, $a\notin\sigma$.
%
%	\item[Freshness]
%Existential variables in the range of $\sigma$ are fresh, i.e.\ distinct from any existentially quantified variable in $\seq AB$.
%\end{description}
\end{definition}

We assume the existential variables occurring in the range of $\Sigma$ to be fresh, i.e.\ distinct from any existentially quantified variable in $\seq AB$.  We abbreviate by $a\in \sigma$ the statement that a variable $a$ occurs free in the range of $\sigma$, i.e.\ that $a\in\fv(\sigma(x))$ for some $x$.

@@

A witness link $\link PQ$ on two atomic formulas $P$ and $Q$ is an \defn{axiom} link if $P\sigma=\dual Q\sigma$; a witness \defn{axiom} linking is one where every link is an axiom link. Observe that the de-sequentialization $[\pi]$ is a witness pre-net with axiom linking $\deseq[\varnothing]\pi AB$.

@@

\noindent
A witness pre-net $\net{\lambda_\Sigma}AB$ \defn{strict-coalesces} if it reduces to a single link $\net{\{\link[\varnothing]AB\}}AB$.\end{definition}

To be explicit, to generate coalescence as a rewrite relation means the following: a pre-net coalesces to another, $\net{\lambda_\Sigma} AB\scoal \net{{\lambda'_{\Sigma'}}}AB$ if and only if there is a rewrite rule
\[
	\net{\kappa_\Xi} CD\scoal\net{\kappa'_{\Xi'}} CD
\] 
such that 
\[
	C\leq A \qquad
	D\leq B \qquad
	\kappa_\Xi\subseteq\lambda_\Sigma \qquad
	\kappa'_{\Xi'}\subseteq\lambda'_{\Sigma'} \quad
	\text{and} \quad
	\lambda_\Sigma-\kappa_\Xi=\lambda'_{\Sigma'}-\kappa'_{\Xi'}~.
\]

Two useful properties of witness nets are preserved over a coalescence sequence, captured in the lemma below. The first property is formulated the way it is to account for vacuous quantifiers, for which no actual variable occurs in $C$ or $D$, as well as formulas such as $\exists x.C\+C'$ where $x$ occurs in $C'$ but not $C$, where we still would like an explicit witness for $x$ on a link $\link CD$.

\begin{lemma}
Every pre-net on a coalescence path $\net{\lambda_\Sigma}AB \scoal^* \net{\{\link[\varnothing]AB^\pi\}}AB$ from a witness net satisfies the following properties. 
\begin{description}
	\item[Eigenvariables not free]
For any link $\link CD$, a universal variable $a\in\varA$ quantified in $C$ or $D$ does not occur free in the range of $\sigma$, $a\notin\sigma$.

	\item[Exact coverage]
For any link $\link CD$, the domain of $\sigma$ is exactly the variables existentially quantified in $\seq AB$ that will be bound in $C$ and $D$, if they occur:
\[
	\dom\sigma = \{\,x \mid C < \exists x.X \leq A\,\}
			\cup \{\,y \mid D < \exists y.Y \leq B\,\}~.
\]
\end{description}
\end{lemma}

\begin{proof}
By induction on the coalescence path, going forward from $\net{\lambda_\Sigma}AB$ for the first property, and working backward from $\net{\{\link[\varnothing]AB^\pi\}}AB$ for the second.
\end{proof}

\begin{theorem}
For any \all1 proof $\pi$, the de-sequentialization $[\pi]$ is a witness net that sequentializes to $\pi$. 
\end{theorem}

\begin{proof}
It follows by induction on $\prf\pi AB$ that if $\lambda_\Sigma=[\pi]_{A',B'}^\sigma$ where $A'\sigma=A$ and $B'\sigma=B$, then $\net{\lambda^\star_\Sigma}AB$ reduces in $(\scoal)$ to $\net{\{\link{A'}{B'}^\pi\}}AB$. The statement is the specific case $\sigma=\varnothing$.
\end{proof}

A formula $\exists x.A$ has two natural immediate descendants: one \emph{syntactic}, its subformula $A$, and one \emph{semantic}, the premise $A[t/x]$ for its introduction rule, for any given $t$. We will define the latter as a tree-order generated from a formula $A$: the \defn{positions} of $A$ are its subformulas under substitution,
\[
	\poss A~=~\{~B'\sigma~\mid~A=A'\sigma~,~B'\sub A'~,~\sigma\in\varE\rightharpoonup\terms~\}
\]

The notion of a \defn{slice} is central to additive proof theory. {\color{red}[TO DO: say what it is]} In the propositional case, slices are readily defined as sets of subformulas. Here, in the first-order case, we will work analogously on the set of positions.

The immediate problem is that the number of positions immediately below $\exists x.B$ is infinite, as there is one immediate descendant $B[t/x]$ for every term $t$. We reduce it to finite width by restricting a slice to just the terms assigned to $x$ by the witness labelling $\Sigma$ in a pre-net $\net{\lambda_\Sigma}AB$.

For a witness linking $\lambda_\Sigma$ and an existential variable $x$, write $\Sigma(x)\subseteq\terms$ for the \defn{witness set} of $x$, which collects the terms assigned to $x$ by the witness maps in $\Sigma$ for every link in $\lambda$:
\[
	\Sigma(x) = \{~\sigma(x)~\mid~\link CD\in\lambda_\Sigma~\}~.
\]
%
Given a witness linking $\lambda_\Sigma$, a \defn{slice} $S$ of a formula $A$ is a set of positions of $A$ given by: $S=\{A\}\cup S'$ where $S'$ is as follows.

\begin{itemize}
	\item
\makebox[70pt][l]{If $A = P$}			then $S'=\varnothing$.
	\item
\makebox[70pt][l]{If $A = B\+C$} 		then $S'=S_B\uplus S_C$ where $S_B$ is a slice of $B$ and $S_C$ a slice of $C$.
	\item
\makebox[70pt][l]{If $A = B\*C$}		then $S'$ is a slice of $B$ or a slice of $C$.
	\item
\makebox[70pt][l]{If $A = \exists x.B$}	then $S'=\uplus_{t\,\in\,\Sigma(x)}S_t$ where each $S_t$ is a slice of $B[t/x]$.
	\item
\makebox[70pt][l]{If $A = \forall a.B$}	then $S'$ is a slice of $B$.
\end{itemize}
% Note: careful with multiple identical terms t! It should always be safe to have more...

\noindent
A \defn{slice} of a sequent $\seq AB$ is the product $S_A\times S_B$ of a slice of each formula. A \defn{slice} of a witness pre-net $\net{\lambda_\Sigma} AB$ over a slice $S_A\times S_B$ of $\seq AB$ is the set of links
\[
	\{~\link CD\in \lambda_\Sigma~\mid~(C\sigma,D\sigma)\in S_A\times S_B~\}~.
\]

Within a slice, we need to ensure that the order in which quantifiers are instantiated is sound. That is, we need to have an analogue of the \emph{eigenvariable condition} on the $\qrr!$-rule of sequent calculus, that the quantified variable $a$ of the primary formula $\forall a.A$ does not occur free in the context. We will define a \defn{dependency} condition to this effect.

Rather than define the condition on slices, since we will require that each slice contains a single link, we will define it for individual links. It is expected that for a more involved logic that includes multiplicative connectives, a dependency condition such as the below one needs to be applied to a slice instead.

In a pre-net $\net{\lambda_\Sigma}AB$, let the \defn{column} of a link $\link CD$ be the set of positions of $A\sigma$ and $B\sigma$ that have $C\sigma$ respectively $D\sigma$ as a descendant,
\[
	\{~X~\mid~C\sigma\sub X\sub A\sigma~\}\cup\{~Y~\mid~D\sigma\leq Y\leq B\sigma~\}~,
\]
with an order $(\dep)$ given by: $X\dep Y$ if $X\leq Y$, and $\exists x.X\dep \forall a.Y$ if $a\in\fv(\sigma(x))$. We will have $X\dep\forall a.Y$ whenever $a$ is free in $X$, plus when $X$ occurs under a vacuous existential quantifier $\exists x$ with $a\in\fv(\sigma(x))$.

\begin{definition}
A witness pre-net is \defn{correct} if it is
\begin{enumerate}
\item \defn{slice-correct}: every slice is a singleton, and
\item \defn{dependency-correct}: every column is a partial order (i.e.\ is acyclic/antisymmetric).
\end{enumerate}
\end{definition}

\begin{lemma}
\label{lem:preservation}
Strict coalescence preserves and reflects correctness: for a strict coalescence step $L\scoal R$, the witness pre-net $L$ is correct if and only if $R$ is.
\end{lemma}

\begin{proof}
Let $L=\net{\lambda_\Sigma}AB$ and $R=\net{\kappa_\Xi}AB$. We will consider the four coalescence steps.
%
\begin{align}
  	\net{\{\link C{D_i}\}}C{D_0{\+}D_1} 
 &~\scoal~
	\net{\{\link C{D_0{\+}D_1} \}}C{D_0{\+}D_1}
\tag{$\srr+i$}
\\	\net{\{\link C{D_0},\link C{D_1}\}}C{D_0{\*}D_1}
 &~\scoal~ 
	\net{\{\link C{D_0{\*}D_1}\}}C{D_0{\*}D_1}
\tag{$\srr*$}
\\	\net{\{\link CD\}}C{\exists x.D}
 &~\scoal~ 
	\net{\{\link[\sigma\minus x]C{\exists x.D}\}}C{\exists x.D}
\tag{$\srr?$}
\\	\net{\{\link CD\}}C{\forall a.D}
 &~\scoal~
	\net{\{\link C{\forall a.D}\}}C{\forall a.D}
\tag{$\srr!$}
\end{align}
%
\begin{description}
	\item[$(\srr+i)$]
Let $S_A$  contain $C\sigma$ and $S_B$ contain $D_0\sigma\+D_1\sigma$. Then $S_B$ contains also $D_0\sigma$ and $D_1\sigma$. If the slice of $L$ over $S_A$ and $S_B$ is $\{\link C{D_i}\}$, then that of $R$ over $S_A$ and $S_B$ is $\{\link C{D_0{\+}D_1}\}$, and vice versa. Since other slices are unaffected, $L$ is slice-correct if and only if $R$ is. For dependency-correctness, the column of $\link C{D_i}$ is that of $\link C{D_0{\+}D_1}$ plus $D_i\sigma$ itself, which is minimal in the order $\dep$.

	\item[$(\srr*)$]
Let $S_A$ contain $C\sigma$, and $S_B$ contain $D_0\sigma\+D_1\sigma$ and either $D_0\sigma$ or $D_1\sigma$. Then the slice of $L$ over $S_A$ and $S_B$ is $\{\link C{D_0}\}$ respectively $\{\link C{D_1}\}$ if and only if that of $R$ is $\{\link C{D_0{\*}D_1}\}$. Dependency-correctness is immediate, as above.

	\item[$(\srr?)$]
The witness sets $\Sigma(x)$ and $\Xi(x)$ for $L$ and $R$ need not be the same, since $L$ has $\sigma$ where $R$ has $\sigma\minus x$. Let $\Sigma(x)=\{t,t_1,\dots,t_n\}$ and $\Xi(x)=\{t_1,\dots,t_n\}$, where $\sigma(x)=t$. Let $S_A$ contain $C\sigma$. Let $S_B$ be a slice of $B$ for $\lambda_\Sigma$  and let $S'_B$ be a corresponding slice of $B$ for $\lambda_\Xi$, both containing $(\exists x.D)\sigma=\exists x.D(\sigma\minus x)$. Then $S_B$ contains $D[t/x],D[t_1/x],\dots,D[t_n/x]$ where $S'_B$ has $D[t_1/x],\dots,D[t_n/x]$. It follows that the slice of $L$ over $S_A$ and $S_B$ is $\{\link CD\}$ if and only if that of $R$ over $S_A$ and $S'_B$ is $\{\link[\sigma\minus x]C{\exists x.D}\}$.
% Note that if $t=t_i$ for some $i$ then still the slice over $C$ and $D[t_i/x]$ in $L$ must be empty.

The column of $\link CD$ is that of $\link[\sigma\minus x]C{\exists x.D}$ plus the formula $D\sigma$ itself. In addition, in the former the order $(\dep)$ is augmented with $(\exists x.D)\sigma\dep\forall a.X$ whenever $a\in\fv(\sigma(x))$. Since $D\sigma$ is the only element smaller than $(\exists x.D)\sigma$, this may only create a cycle in $(\dep)$ if $D=\forall b.Y$; but in that case, $b$ is not free in $\sigma$ (by the ``Eigenvariables not free'' condition on pre-nets), and so cannot create a cycle. It follows that $L$ is dependency-correct if and only if $R$ is.

	\item[$(\srr!)$]
Let $S_A$ contain $C\sigma$, and $S_B$ contain $\forall a.D\sigma$ and $D\sigma$. The slice of $L$ over $S_A$ and $S_B$ is $\{\link CD\}$ if and only if that of $R$ is $\{\link C{\forall a.D}\}$. Dependency-correctness is immediate by the observation that $D\sigma$ is minimal in the column of $\link CD$.

\qedhere
\end{description}
\end{proof}

\begin{lemma}
A witness pre-net $\net{\lambda_\Sigma}AB$ is correct if and only if it strict-coalesces to $\net{\{\link{A'}{B'}\}}AB$ where $A=A'\sigma$ and $B=B'\sigma$. 
\end{lemma}

\begin{proof}
From right to left, we proceed by induction on the coalescence path, with the end result as the base case. It is slice-correct: every slice of $\seq AB$ contains $(A,B)$, and since $(A,B)=(A'\sigma,B'\sigma)$ every slice of $\net{\{\link{A'}{B'}\}}AB$ is the singleton $\{\link{A'}{B'}\}$. It is also dependency-correct: the column of $\link{A'}{B'}$ is the set $\{A,B\}$, where $A$ and $B$ are unrelated in $(\dep)$ (note that even if one is $\exists x.C$, still $x$ is not in the domain of $\sigma$, which covers the existential variables quantified by successors of $A'$ and $B'$ but not $A'$ and $B'$ themselves). By Lemma~\ref{lem:preservation}, coalescence reflects correctness, so that any pre-net along the coalescence path is correct, in particular $\net{\lambda_\Sigma}AB$.

From left to right, let $\net{\lambda_\Sigma}AB$ be correct. Let the \defn{depth} of a link $\link CD$ be the distance from $C$ and $D$ to the root of $A$ respectively $B$, given by a pair of integers $(n,m)$. We will demonstrate that a link at maximal depth in the product order, $(i,j)\leq(n,m)$ if and only if $i\leq n$ and $j\leq m$, may always be coalesced, unless it is the unique link $\link{A'}{B'}$ at $(0,0)$. By Lemma~\ref{lem:preservation} the result of this coalescence step is again correct, and it follows by induction on the decreasing depth of links that the pre-net strict-coalesces to $\net{\{\link{A'}{B'}\}}AB$.

To see that a maximally deep link coalesces, first note that a link $\link C{D_i}$ where $D_i$ occurs in $D_0{\+}D_1$ may always contract, as may a link $\link CD$ where $D$ occurs in $\exists x.D$. This leaves the following cases:
\begin{itemize}
	\item $\link A{D_i}$ with $D_i$ in $D=D_0\*D_1$; without loss of generality, let $i=0$. A slice $S_B$ of $B$ containing $D$ also contains either $D_0$ or $D_1$. If the former, the corresponding slice of $\net{\lambda_\Sigma}AB$ is $\{\link A{D_0}\}$.

	\item  $\link{C_i}{D_j}$ in $C_0\*C_1$ and $D_0\*D_j$

\end{itemize}

@@

We define the following further operations on witness maps.
\[
\begin{tabular}{@{}lp{.75\textwidth}@{}}
	$\sigma\gen\tau$
&
	A witness map is \defn{more general} than another, 
	$\sigma\gen\tau$, if there is a map $\rho$ such that 
	$\sigma\rho=\tau$.
\\[5pt]	
	$\sigma\coh\tau$
&
	Two witness maps are \defn{coherent}, $\sigma\coh\tau$,
	if there is a map $\rho$ such that $\sigma\rho=\tau\rho$.
\\[5pt]
	$\sigma\join\tau$
&
	The \defn{join} of two witness maps $\sigma\join\tau$ is defined
	if they are coherent, and is the least map $\rho$ such that
	$\sigma\gen\rho$ and $\tau\gen\rho$.
\end{tabular}%
\]

To an axiom link $(P,Q)$ we assign an \defn{initial witness map} $\sigma=\init PQ$ as the least witness map in $(\gen)$ such that $P\sigma=\dual Q\sigma$. In other words, $\init PQ$ is the most general unifier ($\textsc{mgu}$) of $P$ and $\dual Q$. An axiom linking $\lambda$ with the initial witness map assigned to each link will be written $\lambda_\star$:
\[
	\lambda_\star = \{~\link PQ~\mid~(P,Q)\in\lambda~,~\sigma=\init PQ~\}~.
\]
For two witness labellings $\Sigma$ and $\Xi$ on the same linking $\lambda$, one is \defn{more general} than the other, $\Sigma\gen\Xi$, if $\Sigma(C,D)\gen\Xi(C,D)$ for every link $(C,D)\in\lambda$. We extend this to the linkings themselves: $\lambda_\Sigma\gen\lambda_\Xi$. Observe that for an axiom linking $\lambda$, the initial linking $\lambda_\star$ is the most general among the witness linkings $\lambda_\Sigma$ for any $\Sigma$.

@@

\\ \\
    \vc{\infer{\crule[\sigma\minus x]A{\exists x.B}}{\crule AB}}
&
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\exists x.B}
    	\Vlink[red,label={~$\scriptstyle\sigma$}]{1,4}
    \end{tikzpicture}}
&\rightsquigarrow&
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\exists x.B}
    	\Vlink[red,label={~$\scriptstyle{\sigma\minus x}$}]{1,1}
    \end{tikzpicture}}
\\ \\
    \vc{\infer[\!\!\scriptstyle{a\,\notin\,\fv(A\sigma)}]{\crule A{\forall a.B}}{\crule AB}}
&    
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\forall a.B}
    	\Vlink[red,label={~$\scriptstyle\sigma$}]{1,4}
    \end{tikzpicture}}
&\stackrel{a\,\notin\,\fv(A\sigma)}\rightsquigarrow&
    \vc{\begin{tikzpicture}[net]
    	\formula[y=2]{A}
    	\formula[y=1]{\forall a.B}
    	\Vlink[red,label={~$\scriptstyle\sigma$}]{1,1}
    \end{tikzpicture}}
\end{array}
\]
\par\bigskip\hrule
\caption{Coalescence derivation rules}
\label{fig:coalescence}
\end{figure}
